How many researchers use 'boilerplate' statistical analysis sections?

### 2021-03-08 manuscript/draft_manuscript.Rmd
add results to manuscript

---
title: "How many researchers use 'boilerplate' statistical analysis sections?"
author: "Nicole White, Richi Nayak, Adrian Barnett"
date: "21/05/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
link-citations: yes
bibliography: references.bib
subtitle: An observational study of papers published in _PLOS ONE_.
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(pander)
library(ggplot2)
panderOptions('table.emphasize.rownames', FALSE)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', Inf)
panderOptions('big.mark', ',')
panderOptions('table.emphasize.rownames', FALSE)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', Inf)
panderOptions('big.mark', ',')
panderOptions('keep.line.breaks', TRUE)
panderOptions('table.alignment.default','left')
```

@@

<!-- # Results -->

<!-- ```{r source, message=F,echo=F} -->
<!-- load('data/search_results_n.RData') -->
<!-- ``` -->

<!-- * total records found: 178654; 131847 unique DOIs -->
<!-- * total stats sections after filtering: 111731 (85%) -->
<!-- * Plot over time. Currently double counting if more than one search query for same doi. -->

<!-- ```{r echo=F,fig.cap="\\label{fig:search-results-n}Figure caption here"} -->
<!-- plot_dat_n = n_by_searchterm_volume %>% group_by(volume) %>%  -->
<!--   summarise(n_matches = sum(n_unique)) %>% right_join(.,n_by_volume %>% select(-n_records),by='volume') %>%  -->
<!--   gather(variable,value,-volume) -->

<!-- plot_dat_n %>% ggplot(.,aes(x=volume,y=value,group=variable))+ -->
<!--   geom_bar(stat='identity',position='dodge',aes(fill=variable)) -->

<!-- ``` -->

@@

* Final sample size
* Breakdown of included articles by year of publication (did/did not include stats section). Final sample broken down by subject classification(s) [most are biology and life sciences]
* Figure showing change in gof measure as function of number of clusters (use to justify choice of number of clusters)

* Desribe topic clusters + examples

>>>>

---
title: "How many researchers use 'boilerplate' statistical analysis sections?"
author: "Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett"
date: "21/05/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
link-citations: yes
bibliography: references.bib
subtitle: An observational study of papers published in _PLOS ONE_.
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(pander)
library(ggplot2)
library(ggpubr)
panderOptions('table.emphasize.rownames', FALSE)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', Inf)
panderOptions('big.mark', ',')
panderOptions('table.emphasize.rownames', FALSE)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', Inf)
panderOptions('big.mark', ',')
panderOptions('keep.line.breaks', TRUE)
panderOptions('table.alignment.default','left')

g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal',
                           axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
```

@@

@@

```{r}
load('total.records.plos.rda')
n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
n.include = total.records[[1]] %>% filter(stage=='Included in analysis') %>% pull(n) %>% sum()
perc.include = round(100*n.include/n.api)
min.perc.stats = total.records[[2]] %>% filter(percentage==min(percentage)) %>% select(volume,percentage)
max.perc.stats = total.records[[2]] %>% filter(percentage==max(percentage)) %>% select(volume,percentage)

```

```{r}
load('wordcount.plos.rda')
n.words = wordcounts %>% summarise(med = median(words),q1=quantile(words,.25),q3=quantile(words,.75))
```

```{r,fig.width=10,fig.height=12,fig.cap='\\label{fig:plos-n}Search results by PLOS ONE volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
to_plot = total.records[[2]] %>% select(-percentage) %>% gather(variable,value,-volume) %>%
  mutate(variable=factor(variable,levels=c('API','Included in analysis')))

g1 = ggplot(to_plot,aes(x=volume,y=value,fill=variable))+geom_bar(stat='identity',position='dodge',width=0.9,colour='black')+scale_y_continuous('Total records',breaks=seq(0,20000,2500))+
  scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+g.theme+theme(legend.title = element_blank())

g2 = ggplot(wordcounts,aes(y=log10(words),x=volume,group=volume))+geom_boxplot()+
  scale_x_continuous('Volume',breaks=1:14)+scale_y_continuous('Word count (log10 transformed)',breaks=seq(0,4,0.5))+g.theme

load('stats_section_subject.classifications.rda')
stats_section_subject = mutate(stats_section_subject,
top3 = ifelse(subject_level_number %in% 1:3,1,0))
to_plot = stats_section_subject %>% group_by(subject_level_name) %>% summarise(n=n(),Yes=sum(top3),No=sum(top3==0),.groups='drop')

to_plot.bar = to_plot %>% arrange(-n) %>% select(-n) %>% gather('Top 3 subject classification',value,-subject_level_name) %>%
mutate(subject_level_name=factor(subject_level_name,levels=rev(unique(subject_level_name))),
'Top 3 subject classification'=factor(`Top 3 subject classification`,levels = c('No','Yes')))

g3 = ggplot(to_plot.bar,aes(x=subject_level_name,y=value/10000,fill=`Top 3 subject classification`))+geom_bar(stat='identity',colour='black',width=0.75)+scale_y_continuous('Total records (x 10,000)',breaks=seq(0,12,1))+xlab('')+coord_flip()+scale_fill_manual(values = c('grey75','grey15'))+ theme_bw() + theme(legend.position = 'top',
legend.direction = 'horizontal',axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))

ggarrange(g1,g2,g3,nrow=3,labels=LETTERS[1:3])

n.top3 = to_plot %>% filter(n==Yes)
n.biology = to_plot %>% filter(subject_level_name=='Biology and life sciences') %>% pull(n)
n.earth = to_plot %>% filter(subject_level_name=='Earth sciences') %>% pull(n)
n.computer = to_plot %>% filter(subject_level_name=='Computer and information sciences') %>% pull(n)

```

API searches returned `r n.api` unique records, of which `r n.include` (`r perc.include`%) included a statistical methods section based on our search criteria. Search results varied by journal volume (Figure 1A). The total number of API search results peaked at volumes 8 (n = `r total.records[[2]] %>% filter(volume==8) %>% pull(API)`) and 9 (n = `r total.records[[2]] %>% filter(volume==8) %>% pull(API)`), corresponding to years 2013 and 2014. This trend aligned with the total number of papers published in PLOS ONE over the same timeframe. The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).

The median length of statistical methods sections after text cleaning was `r n.words[['med']]` per DOI (IQR: `r n.words[['q1']]` to `r n.words[['q3']]`) (Figure 1B).

Across all records included in analysis, all included Biology and life sciences (n = `r n.biology`), Earth sciences
(n = `r n.earth`) and/or Computer and information sciences (n = `r n.computer`) within their top 3 subject classifications
(Figure 1C).
